\texttt{davos}: a Python package ``smuggler'' for constructing lightweight reproducible notebooks

### 2022-10-07 paper/main.tex
working on adressing PF comments on section 1

  Reproducibility is a core requirement of modern scientific research.
  For computational research, reproducibility means that code should
  produce the same results, even when run on different systems.  A
  standard approach to ensuring reproducibility entails packaging a
  project's dependencies along with its primary code base.  Existing
  solutions vary in how deeply these dependencies are specified,
  ranging from virtual environments (which specify all Python package
  versions), to containers (which also specify the operating system),
  to virtual machines (which also specify hardware layers of the
  system).  Each of these existing solutions requires installing or
  setting up a system for running the desired code that must be
  packaged alongside the primary code base.  Here we propose a
  lighter-weight solution than virtual environments: the
  \texttt{davos} library.  When used in combination with a
  notebook-based Python project, the \texttt{davos} library provides a
  mechanism for specifying (and automatically installing) the correct
  package versions of the project's dependencies.  This enables researchers to
  share a complete reproducible environment using a single Jupyter
  notebook file.

@@

Code sharing is a core component of the open science movement that has
inspired the development of a full ecosystem of tools and packages.
However, sharing code, in and of itself, does not guarantee that
others will be able to \textit{reproduce} the desired results.  For
example, research code often requires installing other software
packages that extend the implementation language's basic
functionality.  Within the Python community~\citep{vanR95}, external
packages that are published in the most popular
repositories~\citep{pypi, Anac12} are associated with version numbers
and tags that enable users to guarantee that they are installing
exactly the same code across different computing environments.
Despite that it is \textit{possible} to manually install the intended
version numbers of every dependency of a Python script or package,
doing so may cause conflicts within the user's computing environment
that interfere with the functionality of \textit{other} code.

@@

\subsection{Software architecture}
The \texttt{davos} package consists of two interdependent subpackages. The first comprises a set of ``core" modules that implement\comment{provide?} much of \texttt{davos}'s major functionality, including machinery for installing and validating packages, custom parsers for the \texttt{smuggle} statement (see Section \ref{subsec:smuggle}) and onion comment (see Section \ref{subsec:onion}), and a runtime interface for configuring \texttt{davos}'s behavior (see Section \ref{subsec:config}). However, certain critical\comment{other important} aspects of this functionality require (often substantially) different implementations depending on various properties of the notebook in which \texttt{davos} is used (e.g., whether the frontend is provided by Jupyter or Google Colaboratory, or which version of IPython \cite{PereGran07} is used by the notebook kernel). To deal with this, environment-dependent parts of core features and behaviors are isolated and abstracted to ``helper functions", which are imported from the ``implementations" subpackage. This second subpackage defines multiple, interchangeable implementations of each helper function, organized into per-environment modules. At runtime, \texttt{davos}

>>>>

  Reproducibility is a core requirement of modern scientific research.
  For computational research, reproducibility means that code should
  produce the same results, even when run on different systems.  A
  standard approach to ensuring reproducibility entails packaging a
  project's dependencies along with its primary code base.  Existing
  solutions vary in how deeply these dependencies are specified,
  ranging from virtual environments, to containers, to virtual
  machines.  Each of these existing solutions requires installing or
  setting up a system for running the desired code that must be
  packaged alongside the primary code base.  Here we propose a
  lighter-weight solution than virtual environments: the
  \texttt{davos} library.  When used in combination with a
  notebook-based Python project, the \texttt{davos} library provides a
  mechanism for specifying (and automatically installing) the correct
  package versions of the project's dependencies.  The \texttt{davos}
  library also ensures that those versions are in use any time the
  notebook's code is executed.  This enables researchers to share a complete
  reproducible environment using a single Jupyter notebook file.

@@

Modern research projects that rely on software are becoming
increasingly complex.  Even when a particular function or script
appears simple, the code may depend on external libraries (which in
turn may depend on other libraries), particular CPU or GPU instruction sets, and
so on.  Different versions of those external libraries (or processor
instruction sets) may ....

% IN-PROGRESS TEXT------

The same code, executed on different systems or in different computing
environments, may behave differently.  For example, when research code relies
on external libraries, the code's functionality

Reproducability and predictability are fundamental goals of nearly all
scientific undertakings.  For projects that rely on software
components, achieving these goals

% --------

Code sharing is a core component of the open science movement that has
inspired the development of a large ecosystem of tools and packages.
However, sharing code, in and of itself, does not guarantee that
others will be able to \textit{reproduce} the desired results.  For
example, research code often requires installing other software
packages that extend the implementation language's basic
functionality.  Within the Python community~\citep{vanR95}, external
packages that are published in the most popular
repositories~\citep{pypi, Anac12} are associated with version numbers
and tags that enable users to guarantee that they are installing
exactly the same code across different computing environments.
Despite that it is \textit{possible} to manually install the intended
version numbers of every dependency of a Python script or package,
doing so may cause conflicts within the user's computing environment
that interfere with the functionality of \textit{other} code.

@@

\subsection{Software architecture}
The \texttt{davos} package is structured as two sub-packages: a set of ``core" modules that implement...
